---
title: "project3"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
```



# Climate change and temperature anomalies 


If we wanted to study climate change, we can find data on the *Combined Land-Surface Air and Sea-Surface Water Temperature Anomalies* in the Northern Hemisphere at [NASA's Goddard Institute for Space Studies](https://data.giss.nasa.gov/gistemp). The [tabular data of temperature anomalies can be found here](https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.txt)

To define temperature anomalies you need to have a reference, or base, period which NASA clearly states that it is the period between 1951-1980.

Run the code below to load the file:

```{r weather_data, cache=TRUE}

weather <- 
  read_csv("https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.csv", 
           skip = 1, 
           na = "***")

```

Notice that, when using this function, we added two options: `skip` and `na`.

1. The `skip=1` option is there as the real data table only starts in Row 2, so we need to skip one row. 
1. `na = "***"` option informs R how missing observations in the spreadsheet are coded. When looking at the spreadsheet, you can see that missing data is coded as "***". It is best to specify this here, as otherwise some of the data is not recognized as numeric data.

Once the data is loaded, notice that there is a object titled `weather` in the `Environment` panel. If you cannot see the panel (usually on the top-right), go to `Tools` > `Global Options` > `Pane Layout` and tick the checkbox next to `Environment`. Click on the `weather` object, and the dataframe will pop up on a seperate tab. Inspect the dataframe.

For each month and year, the dataframe shows the deviation of temperature from the normal (expected). Further the dataframe is in wide format. 

You have two objectives in this section:

1. Select the year and the twelve month variables from the `weather` dataset. We do not need the others (J-D, D-N, DJF, etc.) for this assignment. Hint: use `select()` function.

1. Convert the dataframe from wide to 'long' format. Hint: use `gather()` or `pivot_longer()` function. Name the new dataframe as `tidyweather`, name the variable containing the name of the month as `month`, and the temperature deviation values as `delta`.


```{r tidyweather}
tidyweather <- weather %>% 
  select(-c("J-D","D-N","DJF","MAM","JJA","SON")) %>% 
  pivot_longer(cols="Jan":"Dec", names_to = "Month", values_to = "delta")

tidyweather

```