---
title: "project2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of movies- IMDB dataset

  
```{r,load_movies, warning=FALSE, message=FALSE, eval=TRUE}

movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)

```


```{r}
skim(movies)
anyDuplicated(movies)

```
No missing values

No duplicated values since anyDuplicated returns zero



Table with the count of movies by genre, ranked in descending order
```{r}
movies %>% 
  group_by(genre) %>% 
  count(sort = TRUE)
```



Table with the average gross earning, average budget and return on budget by genre
```{r}
movies %>% 
  group_by(genre) %>% 
  summarise(
    gross_average = mean(gross),
    budget_average = mean(budget),
    return_on_budget = gross_average/budget_average
  ) %>% 
  arrange(desc(return_on_budget))

```



Table with top 15 directors who have created the highest gross revenue in the box office
```{r}
movies %>%
  group_by(director) %>% 
  summarise(
    total_gross = sum(gross),
    gross_mean = mean(gross),
    gross_median = median(gross),
    gross_std = sqrt(var(gross))
  ) %>% 
  arrange(desc(total_gross)) %>% 
  head(15)
```



Table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed.
```{r}
ratings_dist <- movies %>% 
  group_by(genre) %>% 
  summarise(
    mean_ratings = mean(rating),
    min_ratings = min(rating),
    max_ratings = max(rating),
    SD = sqrt(var(rating))
  )

ratings_dist

ggplot(ratings_dist, aes(x = mean_ratings)) +
  geom_histogram() +theme_bw()

ggplot(ratings_dist, aes(x = min_ratings)) +
  geom_histogram()+theme_bw()

ggplot(ratings_dist, aes(x = max_ratings)) +
  geom_histogram()+theme_bw()

ggplot(ratings_dist, aes(x = SD)) +
  geom_histogram()+theme_bw()

  
  
```


## Movies

```{r, gross_on_fblikes}
ggplot(movies, aes( x = cast_facebook_likes, y = gross)) +
  labs( x = 'FB Likes', y = 'Gross Revenue') +
  geom_point() +
  theme_bw() +
  NULL

cor(movies$cast_facebook_likes, movies$gross)

```
The graph shows little correlation between the likes and gross and we've further shown this by calculating the correlation.


```{r, gross_on_budget}
ggplot(movies, aes( x = budget , y = gross)) +
  labs( x = 'Budget', y = 'Gross Revenue') +
  geom_point() +
  theme_bw() +
  NULL

cor(movies$budget, movies$gross)

```
The graph shows that budget is likely to be a significant indicator of gross. That's backed up by the correlation factor that gives 0.641

  
```{r, gross_on_rating}
ggplot(movies, aes( x = rating , y = gross)) +
  labs( x = 'Rating', y = 'Gross Revenue') +
  geom_point() + 
  facet_wrap(~genre) +
  theme_bw() +
  NULL

```
IMDB ratings look like a good predictor for gross depending on the genre. For instance, action and adventure genres clearly show a positive correlation, while, horror and sci-fi don't.

