---
title: "project2"
output: html_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
```
# Analysis of movies- IMDB dataset

  
```{r,load_movies, warning=FALSE, message=FALSE, eval=TRUE}

movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)

```


```{r}
skim(movies)
anyDuplicated(movies)

```
No missing values

No duplicated values since anyDuplicated returns zero



Table with the count of movies by genre, ranked in descending order
```{r}
movies %>% 
  group_by(genre) %>% 
  count(sort = TRUE)
```



Table with the average gross earning, average budget and return on budget by genre
```{r}
movies %>% 
  group_by(genre) %>% 
  summarise(
    gross_average = mean(gross),
    budget_average = mean(budget),
    return_on_budget = gross_average/budget_average
  ) %>% 
  arrange(desc(return_on_budget))

```



Table with top 15 directors who have created the highest gross revenue in the box office
```{r}
movies %>%
  group_by(director) %>% 
  summarise(
    total_gross = sum(gross),
    gross_mean = mean(gross),
    gross_median = median(gross),
    gross_std = sqrt(var(gross))
  ) %>% 
  arrange(desc(total_gross)) %>% 
  head(15)
```
Table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed.
```{r}
ratings_dist <- movies %>% 
  group_by(genre) %>% 
  summarise(
    mean_ratings = mean(rating),
    min_ratings = min(rating),
    max_ratings = max(rating),
    SD = sqrt(var(rating))
  )

ratings_dist

ggplot(ratings_dist, aes(x = mean_ratings)) +
  geom_histogram()

ggplot(ratings_dist, aes(x = min_ratings)) +
  geom_histogram()

ggplot(ratings_dist, aes(x = max_ratings)) +
  geom_histogram()

ggplot(ratings_dist, aes(x = SD)) +
  geom_histogram()

  
  
```